Multi-Instance Methods for Partially Supervised Image Segmentation

### 2013-09-19 main.tex
typographic fine-tuning :)

\chapter{Conclusion and Future Direction}
\section{Summary}
This thesis explored the use of structured prediction methods for semantic
segmentation and object class segmentation of natural images, as an important
step towards general scene understanding.
We use the paradigm of structured prediction, which allows for a principled
integration of context and object relations.
%Current research in computer vision progresses at an incredible speed, with new
%methods and applications emerging constantly. Still, there are large
%differences in the paradigms employed by different researchers. While it is
%generally acknowledged that context is one of the most important cues in
%recognizing objects and interpreting scenes, there is little consensus on how
%contextual information should be modelled, or even learned.
%The structured prediction paradigm allows a principled approach to
%incorporating predictions within an image, and to include context information.
%We investigated the potential of structural methods for the application of
%semantic image segmentation and object recognition, applications to which
%structured methods have been applied successfully before.
We focus on \emph{learning} of structural models and the interaction of
inference and learning in the neighborhood models typically employed for
semantic segmentation.
We presents an open source software implementation of a variety of
popular learning algorithms for structural support vector machines, together
with a thorough evaluation of their properties in the presence of approximate
inference. Our software provides a solid basis for future research into learning,
inference and models for computer vision, by providing extensive examples and
benchmarks.

We demonstrate that effective use of available inference mechanisms enables
exact learning, even in the presence of loops in the underlying
factor graph, leading to state-of-the art results on the MSRC-21 dataset.
We also demonstrate the power of conditional interactions, learning spacial
interactions in a RGB-D setting. Our approach improves upon the state-of-the
art on the NYU V2 benchmark for annotation of semantic structure classes.

We also presented a novel approach for clustering based on information theoretic
principles. Our algorithm improves upon methods from the literature on a wide
range of datasets. This indicates that also in the task of extracting
superpixels, we can hope to achieve better results than the $k$-means based
SLIC algorithm that we used.

@@

\section{Future Directions}
There are several directions for future research that we think would be
interesting to pursue as an extension of the presented results:
\paragraph{Large-Scale Weakly Supervised Object Class Segmentation}
We demonstrated a new method for object class segmentation using only weak
supervision. One of the main advantages of such a method is that it is
potentially able to exploit the large amount of weakly labeled data that
is available on the internet. Using additional training data, and evaluating on
the given, manually annotated data, would be an interesting topic for future
research.
\paragraph{Inference Machines}
Recently \citet{stoyanov2011empirical} started a new trend in structured
prediction, which is sometimes called ``inference machines''. The basic
principle is simple: view the process of prediction with a given inference
procedure as a prediction method and optimize parameters of this method using
empirical risk minimization.  The work of \citet{stoyanov2011empirical}, used
loopy belief propagation as their inference algorithm and the optimization is
carried out simply using gradient descent on the non-convex but differentiable
loss function.  Other recent work in this direction includes
\citet{krahenbuhlparameter}, who used mean-field inference in a fully connected
conditional random field and \citet{jancsarylearning}, who used closed form
inference in a Gaussian CRF\@.
While these algorithms show great promise, their relation to the traditional
approach of structured prediction used in this work is mostly unclear. In particular,
if exact traditional learning is possible in a model, it is unclear how much
can be gained by direct empirical risk minimization.
Little empirical comparison was performed, and we are not aware of theoretical
work in this direction, leaving much room for future investigation of this
direction.
\paragraph{Non-Linear Models}
In this work, we only considered models that are linear in the input features--though
features are highly non-linear in the original input pixels. Allowing non-linear
interactions increases the representational power of a CRF, possibly leading
to better segmentation results.
Kernelization of structural support vector machines is trivial in theory, but
had only limited success in the context of CRFs for image
segmentation~\citep{lucchi2012structured}.
Two major alternatives for non-linear CRFs were proposed in the literature,
conditional neural fields~\citep{peng2009conditional} based on neural networks,
and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
Conditional neural fields have only applied to sequence classification so far,
and extending them to our setting of semantic image segmentation would be very
interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
graphs for image processing, but not for higher-level tasks such as semantic
segmentation. If it is possible to include context in a meaningful way, it
might be possible to address even object-centric tasks such as object class
segmentation with DTFs.

\paragraph{Higher Order Potentials and Latent Variable Models}
While non-linear potentials would allow for more complex interactions between
inputs and labelings, introducing higher order potentials or latent variables
allow the model to express more complex interactions within the output
variables. Possible examples are consistency of larger regions, learning parts
or learning of scene classes and cooccurcences.
In principle, higher order potentials and latent variable models are
equivalent, in that each energy function expressed in either form can be
transformed into an energy function of the other kind. In practice, learning of
higher order potentials for semantic segmentation has received little
attention, while most approaches using latent variables are often limited by
the non-convexity of learning.  It would be interesting to compare current
methods using latent variable and higher order approaches, and see how these
interact with different inference and learning schemes.

\paragraph{Feature Design}
This work mostly focused on learning methods, and less on the input, with
the exception of Chapter~\ref{ch:nyu}, that explores the use of 3D features
for semantic segmentation of indoor scenes. It is clear, however, that
the input feature play an important role in the performance of any system.
Using our approach for exact learning of loopy graphs, it seems to be worthwhile to
revisit the works of \citet{nowozin2010parameter} and \citet{lucchi2011spatial},
that evaluate the impact of input features and piecewise training, and of the importance
of global constraints versus global features, respectively.
In particular the importance of features for pairwise potentials has been
somewhat overlooked, often being reduced to a single constant or contrast
sensitive feature.

>>>>

\chapter{Conclusion and Future Direction}
\section{Summary}
In this thesis, we explored the use of structured prediction methods for semantic
segmentation and object class segmentation of natural images, an important
step towards general scene understanding.
We use the paradigm of structured prediction, which allows for a principled
integration of context and object relations.
%Current research in computer vision progresses at an incredible speed, with new
%methods and applications emerging constantly. Still, there are large
%differences in the paradigms employed by different researchers. While it is
%generally acknowledged that context is one of the most important cues in
%recognizing objects and interpreting scenes, there is little consensus on how
%contextual information should be modelled, or even learned.
%The structured prediction paradigm allows a principled approach to
%incorporating predictions within an image, and to include context information.
%We investigated the potential of structural methods for the application of
%semantic image segmentation and object recognition, applications to which
%structured methods have been applied successfully before.
We focused on \emph{learning} of structural models and the interaction of
inference and learning in the neighborhood models typically employed for
semantic segmentation.
We presented an open source software implementation of a variety of
popular learning algorithms for structural support vector machines, together
with a thorough evaluation of their properties in the presence of approximate
inference. Our software provides a solid basis for future research into learning,
inference and models for computer vision, by providing extensive examples and
benchmarks.

We showed that effective use of available inference mechanisms enables
exact learning, even in the presence of loops in the underlying
factor graph. Our methods achieve competitive performance with similar
 methods on the Pascal VOC 2010 dataset, and yield state-of-the art
 results on the MSRC-21 dataset.
We demonstrated the power of conditional interactions by learning spacial
interactions in an RGB-D setting. Here, our approach improves upon the state-of-the
art on the NYU V2 benchmark for annotation of semantic structure classes.

We also presented a novel approach for clustering based on information theoretic
principles. Our algorithm improves upon methods from the literature on a wide
range of datasets. This indicates that in the task of extracting
superpixels, we can also hope to achieve better results than the $k$-means based
SLIC algorithm that we used.

@@

\section{Future Directions}
There are several directions for future research that we think would be
interesting to pursue as an extension of the presented results:
\paragraph{Large-Scale Weakly Supervised Object Class Segmentation}
We demonstrated a new method for object class segmentation using only weak
supervision. One of the main advantages of such a method is that it is
potentially able to exploit the large amount of weakly labeled data that
is available on the internet. Using additional, weakly labeled training data,
and evaluating on the given, manually annotated data, is therefore a promising
path for improving the presented results.
\paragraph{Inference Machines}
Recently \citet{stoyanov2011empirical} started a new trend in structured
prediction, which is sometimes called ``inference machines''. The basic
principle is simple: the process of prediction \emph{using a given inference
procedure} is viewed as a feed-forward method for prediction, and parameters of this
prediction process are optimized directly using empirical risk minimization. The work of
\citet{stoyanov2011empirical}, used loopy belief propagation as their inference
algorithm and the optimization is carried out simply using gradient descent on
the non-convex but differentiable loss function.  Other recent work in this
direction includes \citet{krahenbuhlparameter}, who used mean-field inference
in a fully connected conditional random field and \citet{jancsarylearning}, who
used closed form inference in a Gaussian CRF\@.
While these algorithms show great promise, their relation to the traditional
approach of structured prediction used in this work is mostly unclear. In particular,
if exact traditional learning is possible in a model, it is uncertain how much
in accuracy and efficiency can be gained by direct empirical risk minimization.
Only limited empirical comparison is available, and we are not aware of theoretical
work in this direction, leaving much room for future investigation.
\paragraph{Non-Linear Models}
In this work, we only considered models that are linear in the input features---though
features are highly non-linear in the original input pixels. Allowing non-linear
interactions increases the representational power of a CRF, possibly leading
to more accurate prediction results.
Kernelization of structural support vector machines is straight-forward in theory, but
had only limited success in the context of CRFs for image
segmentation~\citep{lucchi2012structured}.
Two major alternatives for non-linear CRFs were proposed in the literature,
conditional neural fields~\citep{peng2009conditional} based on neural networks,
and decision tree fields~\citep{nowozin2011decision}, based on decision trees.
Conditional neural fields have only been applied to sequence classification so far,
and extending them to our setting of semantic image segmentation would be very
interesting. Decision tree fields~(DTFs) on the other hand have been applied to loopy
graphs for image processing, but not for higher-level tasks such as semantic
segmentation. If it is possible to include context in a meaningful way, it
might be possible to address even object-centric tasks such as object class
segmentation with DTFs.

\paragraph{Higher Order Potentials and Latent Variable Models}
While non-linear potentials would allow for more complex interactions between
inputs and labelings, introducing higher order potentials or latent variables
allow the model to express more complex interactions within the output
variables~\citep{dann2012pottics, kohli2009robust}. Possible examples are
consistency of larger regions, learning parts or learning of scene classes and
cooccurences.
In principle, higher order potentials and latent variable models are
equivalent, in that each energy function expressed in either form can be
transformed into an energy function of the other kind. In practice, learning of
higher order potentials for semantic segmentation has received little
attention, while approaches using latent variables are often limited by
the non-convexity of learning.  It would be interesting to compare current
methods using latent variable and higher order approaches, and see how these
interact with different inference and learning schemes.

\paragraph{Feature Design}
This work mostly focused on learning methods, and less on the input---with
the exception of Chapter~\ref{ch:nyu}, which explores the use of 3D features
for semantic segmentation of indoor scenes. It is clear, however, that
the input features play an important role in the performance of any system.
Using our approach for exact learning of loopy graphs, it seems to be worthwhile to
revisit the works of \citet{nowozin2010parameter} and \citet{lucchi2011spatial},
that evaluate the impact of input features and piecewise training, and of the importance
of global constraints versus global features, respectively.
In particular the importance of features for pairwise potentials has been
somewhat overlooked in the computer vision literature, often being reduced to a
single constant or contrast sensitive feature.

### 2013-09-19 partial_supervision/paper.tex
typographic fine-tuning :)

A bag is labeled positive if it contains at least one positive example, and
negative otherwise.  During training, only the labels of the training bags, not
of the instances inside the bags, are known.  The goal is to learn a classifier
for unseen bags. 
Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we
assume that bags are simply sets, not multi-sets.  Then a bag is an element of
the power set $2^\mathcal{X}$ and the task is to learn a function
\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}.  \end{equation}
Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and
labels $y_i \in \{-1,+1\}$.  It is assumed that the $f_{MI}$ function stems
from a so-called underlying concept, given by an (unknown) function
$f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
\begin{equation}\eqlabel{multi_instance}
f_{MI}(X)= \max_{x \in X}\ f_{I}(x).
\end{equation}

@@

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lccccccc}
    \toprule
        & SVM-SVR & EMDD & mi-SVM & MI-SVM & MICA & MIK & MIK-I \\
    \cmidrule(rl){2-8}
    Musk1 & 87.9 &84.9 &  87.4 &  77.9     & 84.3 & 88.0& 88.0 \\
    Musk2 & 85.4 &84.8 &  83.6 &  84.3     & 90.5 & 89.3& 85.2 \\
    \bottomrule
    \end{tabularx}
    \caption{Bag level performance of various MIL algorithms on the standard Musk
    datasets. All but MIK provide instance-level labeling.\tablabel{muskacc}}

@@

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccc}
    \toprule
    & \multicolumn2c{Musk1}  & \multicolumn2c{Musk2}  \\
                &accuracy & witness-rate & accuracy & witness-rate  \\
    \cmidrule(rl){2-3}
    \cmidrule(rl){4-5}
    mi-SVM      & 87.4          & 100\%               &  83.6          & 83.9\%\\
    SVM-SVR     & 87.9          & 100\%               &  85.4          & 89.5\%\\
    MIK-I& 88.0          & 99\%                &  85.2          & 62.3\%\\
    \bottomrule
    \end{tabularx}
    \caption{MIL algorithms and the empirical witness rates of the
        classifiers.  \tablabel{musk-witness}}
\end{table}

@@

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}p{208pt}ccc}
    \toprule
                & car & bike & person \\
    \cmidrule(l){2-2}
    \cmidrule(l){3-3}
    \cmidrule(l){4-4}
        This work&   0.30&  0.45&  0.43 \\
        Best strongly supervised approaches~\citep{fulkerson2009class,schulz2011}&   0.72&  0.72&  0.66 \\
    \bottomrule
    \end{tabularx}
    \caption{Pixel-level accuracy on the Graz-02 dataset.
        \tablabel{tab:graz}}
\end{table}

>>>>

A bag is labeled positive if it contains at least one positive example, and
negative otherwise.  During training, only the labels of the training bags, not
of the instances inside the bags, are known.  The goal is to learn a classifier
for unseen bags. 
Formally, let $\mathcal{X}$ be the set of instances. To simplify notation, we
assume that bags are simply sets, not multi-sets.
\pagebreak\\
Then a bag is an element of
the power set $2^\mathcal{X}$ and the task is to learn a function
\begin{equation} f_{MI} \colon 2^\mathcal{X} \rightarrow \{-1,+1\}.  \end{equation}
Training examples are tuples $(X_i,y_i)$ of bags $X_i \subset \mathcal{X}$ and
labels $y_i \in \{-1,+1\}$.  It is assumed that the $f_{MI}$ function stems
from a so-called underlying concept, given by an (unknown) function
$f_{I} \colon \mathcal{X} \rightarrow \{-1,+1\}$, with 
\begin{equation}\eqlabel{multi_instance}
f_{MI}(X)= \max_{x \in X}\ f_{I}(x).
\end{equation}

@@

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lccccccc}
    \toprule
        & SVM-SVR & EMDD & mi-SVM & MI-SVM & MICA & MIK & MIK-I \\
    \cmidrule{2-8}
    Musk1 & 87.9 &84.9 &  87.4 &  77.9     & 84.3 & 88.0& 88.0 \\
    Musk2 & 85.4 &84.8 &  83.6 &  84.3     & 90.5 & 89.3& 85.2 \\
    \bottomrule
    \end{tabularx}
    \caption{Bag level performance of various MIL algorithms on the standard Musk
    datasets. All but MIK provide instance-level labeling.\tablabel{muskacc}}

@@

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lcccc}
    \toprule
    & \multicolumn2c{Musk1}  & \multicolumn2c{Musk2}  \\
                &accuracy & witness-rate & accuracy & witness-rate  \\
    \cmidrule{2-3}
    \cmidrule{4-5}
    mi-SVM      & 87.4          & 100\%               &  83.6          & 83.9\%\\
    SVM-SVR     & 87.9          & 100\%               &  85.4          & 89.5\%\\
    MIK-I& 88.0          & 99\%                &  85.2          & 62.3\%\\
    \bottomrule
    \end{tabularx}
    \caption{MIL algorithms and the empirical witness rates of the
        classifiers.  \tablabel{musk-witness}}
\end{table}

@@

\begin{table}
    \centering
    \begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}p{208pt}ccc}
    \toprule
                & car & bike & person \\
    \cmidrule{2-4}
        This work&   0.30&  0.45&  0.43 \\
        Best strongly supervised approaches~\citep{fulkerson2009class,schulz2011}&   0.72&  0.72&  0.66 \\
    \bottomrule
    \end{tabularx}
    \caption{Pixel-level accuracy on the Graz-02 dataset.
        \tablabel{tab:graz}}
\end{table}
